---
title: "Chapter 1"
author: "Roxana Agrigoroaie"
date: "24 05 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Joining Tables

LEGO dataset: made up of multiple tables

```{r}
# a table with the lego sets throughout time
sets <- readRDS("data/sets.rds")
glimpse(sets)
```

There is a theme_id, which links the sets table to the themes table.

```{r}
themes <- readRDS("data/themes.rds")
themes
```

To see the theme that each set is associated with a join is needed: inner_join.

```{r}
sets %>%
  inner_join(themes, by = c("theme_id" = "id"))
```

As both tables have the variable "name", we get name.x for the sets name, and name.y for the themes name

This can be customized using the inner_join

```{r}
sets %>%
  inner_join(themes, by = c("theme_id" = "id"), suffix = c("_set", "_theme"))
```

Most common themes:

```{r}
sets %>%
  inner_join(themes, by = c("theme_id" = "id"), suffix = c("_set", "_theme")) %>%
  count(name_theme, sort = TRUE)
```

Other LEGO tables: parts and part_categories.

Part = shape or figurine

```{r}
parts <- readRDS("data/parts.rds")
part_categories <- readRDS("data/part_categories.rds")
```

```{r}
glimpse(parts)
```

```{r}
glimpse(part_categories)
```

Exercise 1:  Join a list of LEGO parts, available as parts, with these parts' corresponding categories, available as part_categories.

```{r}
parts %>%
  inner_join(part_categories, by = c("part_cat_id" = "id"), suffix = c("_part", "_category"))
```

Joining with a one-to-many relationship

The inventories table: product made up of combination of parts

Joining sets and inventories

```{r}
inventories <- readRDS("data/inventories.rds")
```

```{r}
glimpse(inventories)
```

```{r}
glimpse(sets)
```

In both tables I have the same column: set_num. Therefore I do not need to specify explicitly the name of the column for each table.

```{r}
sets %>%
  inner_join(inventories, by = "set_num")
```

Filtering the joined table

```{r}
sets %>%
  inner_join(inventories, by = "set_num") %>%
  filter(version == 1)
```

Inventory parts -> combines a part with a color 

```{r}
inventory_parts <- readRDS("data/inventory_parts.rds")
glimpse(inventory_parts)
```

Exercise 2: Connect the parts and inventory_parts tables by their part numbers using an inner join.

```{r}
glimpse(parts)
```

```{r}
parts %>%
  inner_join(inventory_parts, by = "part_num")
```

The number of rows went from almost 5000 to more than 250000 (due to the one-to-many relationship).

An inner_join works the same way with either table in either position.

Exercise 3: Connect the inventory_parts and parts tables by their part numbers using an inner join.

```{r}
inventory_parts %>%
  inner_join(parts, by = "part_num")
```

Joining three or more tables

```{r}
sets %>%
  inner_join(inventories, by = "set_num") %>%
  inner_join(themes, by = c("theme_id" = "id"), suffix = c("_set", "_theme"))
```

Exercise 4: Combine the inventories table with the sets table.
Next, join the inventory_parts table to the table you created in the previous join by the inventory IDs.

```{r}
glimpse(sets)
```

```{r}
glimpse(inventories)
```

```{r}
glimpse(inventory_parts)
```

```{r}
sets %>%
  inner_join(inventories, by = "set_num") %>%
  inner_join(inventory_parts, by = c("id" = "inventory_id"))
```

What's the most common color?
Exercise 5a: Inner join the colors table using the color_id column from the previous join and the id column from colors; use the suffixes "_set" and "_color".

```{r}
colors <- readRDS("data/colors.rds")
glimpse(colors)
```


```{r}
sets %>%
  inner_join(inventories, by = "set_num") %>%
  inner_join(inventory_parts, by = c("id" = "inventory_id")) %>% 
  inner_join(colors, by = c("color_id" = "id"), suffix = c("_set", "_color"))
  
```

Exercise 5b: Count the name_color column and sort the results so the most prominent colors appear first.

```{r}
sets %>%
  inner_join(inventories, by = "set_num") %>%
  inner_join(inventory_parts, by = c("id" = "inventory_id")) %>% 
  inner_join(colors, by = c("color_id" = "id"), suffix = c("_set", "_color")) %>%
  count(name_color, sort = TRUE)
```

