---
title: "Selecting and transforming data"
author: "Roxana Agrigoroaie"
date: "21 05 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
counties <- readRDS('data/counties.rds')
```

## Chapter 3: Selecting and Transforming data

Select: only keep columns of interest

```{r}
# Select a range of columns
counties %>%
  select(state, county, drive:work_at_home)
```

```{r}
# Select and arrange

counties %>%
  select(state, county, drive:work_at_home) %>%
  arrange(drive)
```

Select helpers


```{r}
# Contains

counties %>%
  select(state, county, contains("work"))

# Starts with
counties %>%
  select(state, county, starts_with("income"))

```

Other select helpers: contains(); starts_with(); ends_with(); last_col();

For more: ?select_helpers

Removing a variable:

```{r}
counties %>%
  select(-census_id)
```

Exercise 1a: Use glimpse() to examine all the variables in the counties table.

```{r}
glimpse(counties)
```


Exercise 1b: Select the columns for state, county, population, and (using a colon) all five of those industry-related variables; there are five consecutive variables in the table related to the industry of people's work: professional, service, office, construction, and production.

Exercise 1c: Arrange the table in descending order of service to find which counties have the highest rates of working in the service industry.


```{r}
counties %>%
  select(state, county, population, professional:production) %>%
  arrange(desc(service))
```

Exercise 2a: Select the columns state, county, population, and all those that end with work.

Exercise 2b: Filter just for the counties where at least 50% of the population is engaged in public work.

```{r}
counties %>%
  select(state, county, population, ends_with("work")) %>%
  filter(public_work > 50)
```

Rename: change the name of a column

```{r}
counties %>%
  rename(unemployment_rate = unemployment)
```


```{r}
counties %>%
  select(state, county, population, unemployment_rate = unemployment)
```

Exercise 3a: Use count() to determine how many counties are in each state.

```{r}
counties %>%
  count(state)
```

Exercise 3b: Notice the n column in the output; use rename() to rename that to num_counties.

```{r}
counties %>%
  count(state) %>%
  rename(num_counties = n)
```

Exercise 4: Select the columns state, county, and poverty from the counties dataset; in the same step, rename the poverty column to poverty_rate.

```{r}
counties %>%
  select(state, county, poverty_rate = poverty)
```

Transmute: like a combination of select and mutate. It returns a subset of columns that are transformed and changed. 

```{r}
counties %>%
  transmute(state, county, fraction_men = men / population)
```

```{r}
counties %>%
  transmute(state, county, fraction_men = men / population, fraction_women = women / population)
```


Exercise 5: Keep only the state, county, and population columns, and add a new column, density, that contains the population per land_area. Filter for only counties with a population greater than one million. Sort the table in ascending order of density.

```{r}
counties %>%
  transmute(state, county, population, density = population / land_area) %>%
  filter(population > 1000000) %>%
  arrange(density)
```

Exercise 6a: Choose the right verb for changing the name of the unemployment column to unemployment_rate. 

```{r}
counties %>%
  rename(unemployment_rate = unemployment)
```

Exercise 6b: Choose the right verb for keeping only the columns state, county, and the ones containing poverty.

```{r}
counties %>%
  select(state, county, contains("poverty"))
```

Exercise 6c: Calculate a new column called fraction_women with the fraction of the population made up of women, without dropping any columns.

```{r}
counties %>%
  mutate(fraction_women = women / population)
```


Exercise 6d: Keep only three columns: the state, county, and employed / population, which you'll call employment_rate.

```{r}
counties %>%
  transmute(state, county, employment_rate = employed / population)
```

