---
title: 'Chapter 4: Case study'
author: "Roxana Agrigoroaie"
date: "23 05 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
babynames <- readRDS("data/babynames.rds")
```

## Case study: The babynames dataset

```{r}
glimpse(babynames)
```

Frequency of a name

```{r}
babynames %>%
  filter(name == "Amy")
```

```{r}
babynames %>%
  filter(name == "Amy") %>%
  ggplot(aes(x = year, y = number)) + 
  geom_line()
```

Filter for multiple names

```{r}
babynames %>%
  filter(name %in% c("Amy", "Christopher")) %>%
  ggplot(aes(y = number, x = year, color = name)) + 
  geom_line()
```

When was each name most common?

```{r}
babynames %>%
  group_by(name) %>%
  top_n(1, number)
```

Exercise 1: Filter for only the year 1990. Sort the table in descending order of the number of babies born.

```{r}
babynames %>%
  filter(year == 1990) %>%
  arrange(desc(number))
```

Exercise 2: Use group_by() and top_n() to find the most common name for US babies in each year.

```{r}
babynames %>%
  group_by(year) %>%
  top_n(1, number)
```

Exercise 3a: Filter for only the names Steven, Thomas, and Matthew, and assign it to an object called selected_names.

Exercise 3b: Visualize those three names as a line plot over time, with each name represented by a different color.

```{r}
selected_names <- babynames %>%
  filter(name %in% c("Steven", "Thomas", "Matthew"))

ggplot(selected_names, aes(x = year, y = number, color = name)) + 
  geom_line()
```

Grouped mutates: combining group_by() and mutate()

```{r}
babynames %>%
  group_by(year) %>%
  mutate(year_total = sum(number))
```

The result is still grouped, therefore we use ungroup().

Add fraction of people with a specific name in each year.

```{r}
babynames %>%
  group_by(year) %>%
  mutate(year_total = sum(number)) %>%
  ungroup() %>%
  mutate(fraction = number / year_total)
```

Exercise 4a: First, calculate the total number of people born in that year in this dataset as year_total. Next, use year_total to calculate the fraction of people born in each year that have each name.

```{r}
babynames %>%
  group_by(year) %>%
  mutate(year_total = sum(number)) %>%
  ungroup() %>%
  mutate(fraction = number / year_total)
```

Exercise 4b: Now use your newly calculated fraction column, in combination with top_n(), to identify the year each name is most common.

```{r}
babynames %>%
  group_by(year) %>%
  mutate(year_total = sum(number)) %>%
  ungroup() %>%
  mutate(fraction = number / year_total) %>%
  group_by(name) %>%
  top_n(1, fraction)
```

Exercise 5a:  Use a grouped mutate to add two columns: name_total, with the sum of the number of babies born with that name in the entire dataset. name_max, with the maximum number of babies born in any year.

```{r}
babynames %>% 
  group_by(name) %>%
  mutate(name_total = sum(number)) %>%
  mutate(name_max = max(number))
```

Exercise 5b: Add another step to ungroup the table.
Exercise 5c: Add a column called fraction_max containing the number in the year divided by name_max.

```{r}
babynames %>% 
  group_by(name) %>%
  mutate(name_total = sum(number)) %>%
  mutate(name_max = max(number)) %>%
  ungroup() %>%
  mutate(fraction_max = number / name_max)
```


Exercise 6:

```{r}
names_normalized <- babynames %>%
                     group_by(name) %>%
                     mutate(name_total = sum(number),
                            name_max = max(number)) %>%
                     ungroup() %>%
                     mutate(fraction_max = number / name_max)
```

Filter the names_normalized table to limit it to the three names Steven, Thomas, and Matthew.
Create a line plot to visualize fraction_max over time, colored by name.

```{r}
names_normalized %>%
  filter(name %in% c("Steven", "Thomas", "Matthew")) %>%
  ggplot(aes(x = year, y = fraction_max, color = name)) +
  geom_line()
```
Window functions:

takes a vector and return another vector of the same length. 

lag() -> moves each element of a vector to the right

Compare consecutive steps:
vector - lag(vector)

```{r}
babynames_fraction <- babynames %>%
  group_by(year) %>%
  mutate(year_total = sum(number)) %>%
  ungroup() %>%
  mutate(fraction = number / year_total)

babynames_fraction %>%
  filter(name == "Matthew") %>%
  arrange(year) %>%
  mutate(difference = fraction - lag(fraction)) %>%
  # TO get the biggest jump in popularity 
  arrange(desc(difference))
```

Changes within every name

```{r}
babynames_fraction %>%
  arrange(name, year) %>%
  mutate(difference = fraction - lag(fraction)) %>%
  group_by(name) %>%
  arrange(desc(difference))
```

Exercise 7: Arrange the data in ascending order of name and then year. Group by name so that your mutate works within each name. Add a column ratio containing the ratio (not difference) of fraction between each year.

```{r}
babynames_fraction %>%
  arrange(name, year) %>%
  group_by(name) %>%
  mutate(ratio = fraction / lag(fraction))
```


Exercise 8: From each name in the data, keep the observation (the year) with the largest ratio; note the data is already grouped by name.
Sort the ratio column in descending order.
Filter the babynames_ratios_filtered data further by filtering the fraction column to only display results greater than or equal to 0.001.

```{r}
babynames_ratios_filtered <- babynames_fraction %>%
                     arrange(name, year) %>%
                     group_by(name) %>%
                     mutate(ratio = fraction / lag(fraction)) %>%
                     filter(fraction >= 0.00001)
```

```{r}
babynames_ratios_filtered %>%
  top_n(1, ratio) %>%
  arrange(desc(ratio)) %>%
  filter(fraction >= 0.001)
```

