---
title: "Aggregating data"
author: "Roxana Agrigoroaie"
date: "21 05 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Chapter 2:Aggregating data

Count: Find the number of observations

```{r}
counties <- readRDS('data/counties.rds')
```

```{r}
counties %>%
  count()
```

Count variable

```{r}
counties %>%
  count(state) # number of counties in each state
```

Count and sort

Sort the counted data to find the most common observations

```{r}
counties %>%
  count(state, sort = TRUE)
```


Count population

wt => it computes the sum(wt) for each group

```{r}
counties %>%
  count(state, wt = population, sort = TRUE)
```

```{r}
counties_selected <- counties %>%
  select(county, region, state, population, citizens)
```


Exercise 1: Use count() to find the number of counties in each region, using a second argument to sort in descending 

```{r}
counties_selected %>%
  count(region, sort = TRUE)
```

Exercise 2: Count the number of counties in each state, weighted based on the citizens column, and sorted in descending order.

```{r}
counties_selected %>%
  count(state, wt = citizens, sort = TRUE)
```

```{r}
counties_selected <- counties %>%
  select(county, region, state, population, walk)
```

Exercise 3a: Use mutate() to calculate and add a column called population_walk, containing the total number of people who walk to work in a county.

Exercise 3b: Use a (weighted and sorted) count() to find the total number of people who walk to work in each state.

```{r}
counties_selected %>%
  mutate(population_walk = walk * population / 100) %>%
  count(state, wt = population_walk, sort = TRUE)
```

Summarize: takes many observations and turns them into one observation

```{r}
counties %>%
  summarize(total_population = sum(population))
```

Aggregate and summarize

```{r}
counties %>%
  summarize(total_population = sum(population),
            average_unemployment = mean(unemployment))
```

Summary functions:

- sum()

- mean()

- median()

- min()

- max()

- n() -> size of the group

Aggregate within groups: group_by

```{r}
counties %>%
  group_by(state) %>%
  summarize(total_pop = sum(population),
            avg_unemployment = mean(unemployment))
```

Arrange: to sort

```{r}
counties %>%
  group_by(state) %>%
  summarize(total_pop = sum(population),
            avg_unemployment = mean(unemployment)) %>%
  arrange(desc(avg_unemployment))
```


```{r}
counties %>%
  group_by(state, metro) %>%
  summarize(total_pop = sum(population))
```
```{r}
counties_selected <- counties %>%
  select(county, population, income, unemployment)
```


Exercise 4: Summarize the counties dataset to find the following columns: min_population (with the smallest population), max_unemployment (with the maximum unemployment), and average_income (with the mean of the income variable).

```{r}
counties_selected %>%
  summarize(min_population = min(population),
            max_unemployment = max(unemployment),
            average_income = mean(income))
```

```{r}
counties_selected <- counties %>%
  select(state, county, population, land_area)
```

Exercise 5a: Group the data by state, and summarize to create the columns total_area (with total area in square miles) and total_population (with total population).

```{r}
counties_selected %>%
  group_by(state) %>%
  summarize(total_area = sum(land_area),
            total_population = sum(population))
```

Exercise 5b: Add a density column with the people per square mile, then arrange in descending order.

```{r}
counties_selected %>%
  group_by(state) %>%
  summarize(total_area = sum(land_area),
            total_population = sum(population)) %>%
  mutate(density = total_population / total_area) %>%
  arrange(desc(density))
```

```{r}
counties_selected <- counties %>%
  select(region, state, county, population)
```

Exercise 6a: Summarize to find the total population, as a column called total_pop, in each combination of region and state.


```{r}
counties_selected %>%
  group_by(region, state) %>%
  summarize(total_pop = sum(population))
```

Exercise 6b: Notice the tibble is still grouped by region; use another summarize() step to calculate two new columns: the average state population in each region (average_pop) and the median state population in each region (median_pop).

```{r}
counties_selected %>%
  group_by(region, state) %>%
  summarize(total_pop = sum(population)) %>%
  summarize(average_pop = mean(total_pop),
            median_pop = median(total_pop))

```

top_n verb: (superseded now, it is recommended to use slice_min() or slice_max())

```{r}
counties %>%
  select(state, county, population, unemployment, income) %>%
  group_by(state) %>%
  top_n(1, population)
```

```{r}
counties_selected <- counties %>%
  select(region, state, county, metro, population, walk)
```

Exercise 7: Find the county in each region with the highest percentage of citizens who walk to work.

```{r}
counties_selected %>%
  group_by(region) %>%
  top_n(1, walk)
```

```{r}
counties_selected <- counties %>%
  select(region, state, county, population, income)
```


Exercise 8: Calculate the average income (as average_income) of counties within each region and state. Find the highest income state in each region.

```{r}
counties_selected %>%
  group_by(region, state) %>%
  summarize(average_income = mean(income)) %>%
  top_n(1, average_income)
```


```{r}
counties_selected <- counties %>%
  select(state, metro, population)
```

Exercise 9a: For each combination of state and metro, find the total population as total_pop.

Exercise 9b: Extract the most populated row from each state, which will be either Metro or Nonmetro.

Exercise 9c: Ungroup, then count how often Metro or Nonmetro appears to see how many states have more people living in those areas.

```{r}
counties_selected %>%
  group_by(state, metro) %>%
  summarize(total_pop = sum(population)) %>%
  top_n(1, total_pop) %>%
  ungroup() %>%
  count(metro)
```

